Random Fibonacci Anyons Simulable via Covariant Action of Mapping Class Group

### 2014-08-31 ising/Fibonacci/about.tex

%\title{Error Correction in a Phenomenological Fibonacci Anyon Quantum Code}
\title{Random Fibonacci Anyons Simulable via Covariant Action of Mapping Class Group}

\author{Simon Burton, Courtney Brell, Steven Flammia}
%\date{\today}

@@

@@

@@

% ----------------------------------------------------------------------------
%

\heading{Braid group}

Braid group on $n$ strands is the group $B_n$ generated by $n-1$ generators
$\sigma_1, \sigma_2, ... \sigma_{n-1}$ with the following relations:
    $$ \sigma_i \sigma_j = \sigma_j \sigma_i \ \text{when}\ |i-j| > 2, $$
    $$ \sigma_i \sigma_{i+1} \sigma_i =  \sigma_{i+1} \sigma_i \sigma_{i+1}.$$

@@

We define a curve diagram as an embedding $f : L\to D$ that
restricts to a bijection $R_n\to Q_n,$ modulo
isotopy.
This will be denoted as $f : L_n\to D_n.$
We can generalize this to $f : L_m\to D_n$ with $m\leq n$
to mean a map $L\to D$ that restricts to an injection $R_m\to Q_n,$
modulo isotopy.

\begin{center}
\includegraphics[width=0.5\textwidth]{curve-diagram.eps}
\end{center}

@@

The mapping class group $MCG(D_n)$ acts on curve diagrams $f : L_n\to D_n$
via post-composition (left multiplication). See \cite{Dehornoy02}, chapter 6.

@@

In particular, we can single out specific ``braid'' elements of $MCG(D_n)$.
These act on the image of a pair-of-pants embedding: $g:D_2\to D_n.$

@@

@@

\heading{Half-twists}

@@

Each half-twist corresponds to a braid action of two neighbouring
holes on $f$.
See \cite{Kassel10}, Section 1.6.2, for more details on half-twists.

@@

We measure charge contained within a disc. % region, circle
Commuting measurements are either disjoint discs or
one wholly contained within the other.
So (proof?) we get a basis for the Hilbert space from
a pair-of-pants (POP) decomposition:

@@

The feet of the POP form the $n$ holes $Q_n$ in the disc $D$.
\footnote{Every surface that is not a sphere, torus, disc or annulus, has
a POP decomposition, see \cite{Ivanov01}, Theorem 2.4.A.}

@@

Conversely, given a curve $f:L_n\to D_n$
and a binary rooted tree $t$,
we can recover a unique POP that admits $f$.
(proof: use small neighbourhoods of the curve)

@@

@@

We aim discretize the problem for software implementation.
Our first task is to characterize what happens to a curve
if we ``forget'' any braiding inside a disc.
We adapt the Abramsky combinatorial formalism 
of the Temperly-Lieb algebra to our needs (\cite{Abramsky08}, Proposition 1.6).
The following set of ``planar diagrams'' as is defined as:

    $$ T = \{ f(L)\cap E / \sim \ \text{s.t.}\ f:L\to D \}$$

where $f:L\to D$ is a curve (not an isotopy class)
whose image $f(L)$ intersects a disc $E\subset D$ in a finite number of components
and the equivalence relation $\sim$ consists of 
isotopies that fix $\partial E.$ % Local (relative?) isotopies

This object has a simple combinatorial description 
as a string consisting of
the four labels {\bf (, ), H, T} such that 
the parentheses are balanced. The two labels {\bf H} and {\bf T} specify the
head and tail endpoints of $f$, each of which appears at most once in the string.

For example, the following diagram is represented by
the string {\bf(())H}:

@@

%Now we consider a tiling of the disc into subdiscs that are disjoint on
%their interiors.
We consider now a cellulation of a disc.
Concretely, we choose a regular $l$ by $l$ square tiling of a square region.

@@

\important{Definition:} the discrete curve space $C_n^(l)$ consists of
curves $L_n\to D_n$ modulo isotopy that fixes the boundary of
every tile.

>>>>

@@

\title{Random Fibonacci Anyons Simulable via Covariant Action of Mapping Class Group}

\author{Simon Burton, Courtney G. Brell, Steven T. Flammia}

%\date{\today}

%\author{Simon Burton}
%\affiliation{Centre for Engineered Quantum Systems, School of Physics, The University of Sydney, Sydney, Australia}
%
%\author{Courtney G. Brell}
%\affiliation{Institut f\"ur Theoretische Physik, Leibniz Universit\"at Hannover, Appelstra\ss e 2, 30167 Hannover, Germany}
%\affiliation{Centre for Engineered Quantum Systems, School of Physics, The University of Sydney, Sydney, Australia}
%
%\author{Steven T. Flammia}
%\affiliation{Centre for Engineered Quantum Systems, School of Physics, The University of Sydney, Sydney, Australia}

@@

%{\it $^1$ Centre for Engineered Quantum Systems, School of Physics, The University of Sydney, Sydney, Australia}
%{\it $^2$ Institut f\"ur Theoretische Physik, Leibniz Universit\"at Hannover, Appelstra\ss e 2, 30167 Hannover, Germany}

@@

@@

We define a {\it curve diagram} as an embedding $f : L\to D$ that
restricts to a bijection $R_n\to Q_n,$ modulo
isotopy.
This will be denoted as $f : L_n\to D_n.$
We can generalize this to $f : L_m\to D_n$ with $m\leq n$
to mean a map $L\to D$ that restricts to an injection $R_m\to Q_n,$
modulo isotopy.

\begin{center}
\includegraphics[width=0.4\textwidth]{curve-diagram.eps}
\end{center}

@@

The mapping class group $MCG(D_n)$ acts on curve diagrams $f : L_n\to D_n$
via post-composition (left multiplication).
This action is transitive and XXX so we can identify an element of
$MCG(D_n)$ by its action on any curve $f:L_n\to D_n.$
See \cite{Dehornoy02}, chapter 6.

@@

We single out specific ``braid'' elements of $MCG(D_n)$.
These act on the image of a pair-of-pants embedding: $g:D_2\to D_n.$

@@

The {\it braid group} on $n$ strands is the group $B_n$ generated by $n-1$ generators
$\sigma_1, \sigma_2, ... \sigma_{n-1}$ with the following relations:
    $$ \sigma_i \sigma_j = \sigma_j \sigma_i \ \text{when}\ |i-j| > 2, $$
    $$ \sigma_i \sigma_{i+1} \sigma_i =  \sigma_{i+1} \sigma_i \sigma_{i+1}.$$

@@

@@

Each half-twist corresponds to a braid action of two neighbouring
holes on $f$.
\footnote{See \cite{Kassel10}, Section 1.6.2, for more details on half-twists.}

@@

We measure charge contained within a disc. % region, circle
Measurement outcomes belong to a set of charges $\Lambda.$
Commuting measurements are either disjoint discs or
one wholly contained within the other.
So (proof?) we get a basis for the Hilbert space of our system from
a pair-of-pants (POP) decomposition:

@@

The feet of the POP decomposition form the $n$ holes $Q_n$ in the disc $D$.
\footnote{Every surface that is not a sphere, torus, disc or annulus, has
a POP decomposition, see \cite{Ivanov01}, Theorem 2.4.A.}

@@

Conversely, given a curve $f:L_n\to D_n$
and a binary rooted tree $t$,
we can recover a unique POP that admits $f$.
The above POP would be represented by the following tree:

@@

(proof: use small neighbourhoods of the curve)

@@

We aim discretize the problem for software implementation.
Our first task is to characterize what happens to a curve
if we ``forget'' any braiding inside a sub-disc.
We adapt the Abramsky combinatorial description 
of the Temperly-Lieb algebra to our needs
(see \cite{Abramsky08}, section 6.) %Proposition 1.6).

The set of {\it planar diagrams} is defined as:

    $$ P = \{ f(L)\cap E / \sim \ \text{s.t.}\ f:L\to D \}$$

where $f:L\to D$ is a continuous function 
whose image $f(L)$ intersects a disc $E\subset D$ in a finite number of components,
and the equivalence relation $\sim$ consists of 
isotopies that fix $\partial E.$ % Local (relative?) isotopies

This object has a simple combinatorial description 
as a string consisting of
the four labels {\bf (, ), H, T} such that 
the parentheses are balanced.
To construct the string, begin at a point on $\partial E$, and proceed
clockwise. Visiting a strand for the first time produces a {\bf ( } label,
and visiting a strand for the second time produces a {\bf ) } label.
The two labels {\bf H} and {\bf T} specify the
head and tail endpoints of $f$, each of which appears at most once in the string.

For example, the following planar diagram can be represented by
the string {\bf(())H}:

@@

%Now we consider a tiling of the disc into subdiscs that are disjoint on
%their interiors.
We now consider a cellulation of a disc.
Concretely, we choose a regular $l$ by $l$ square tiling of a square region.

@@

\important{Definition:} the discrete curve space $C_n^{(l)}$ consists of
curves $L_n\to D_n$ modulo isotopy that fixes the boundary of
every tile.
